Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-11-07 AnalyticStructure_main.tex
update

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems,
in particular referring to LR-TDDFT calculation. We will investigate the impact that
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
In order to do this, we perform analytic considerations on the equations beyond LR treatment.
\added[id=dal]{This analysis evidences that the description of molecular systems beyond the ground state unavoidably leads to the inclusion of delocalised
degrees of freedom above a specific energetic threshold.}
Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility -
from others that would \emph{explicitly} depend of the employed computational treatment. \added[id=dal]{Moreover on the basis of this analysis we can assess
something about the reliability of the results provided by computational setups able to explicit include the presence of delocalised degrees of freedom compared
to the ones in which these degrees of freedom are excluded by design.}
\deleted[id=dal]{In particular, we will discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects.}

@@

Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of analysing the impact of the BC, it is instructive to transform
the Schr\"odinger equation in a scattering problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \GH(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\GH(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

@@

On the other hand, positive energy solutions behave differently.
Even though the potential is localized, their positive energy are
 spherical waves and the convolution with the source term provides a delocalized wavefunction $\psi(\r)$ in the whole space.
As they cannot be normalized they belong to the essential spectrum of the Hamiltonian operator and form a \emph{continuum} of states.
Outside of the support of the $\ket{\op V \psi}$ the solutions of Eq.~\eqref{SEHelmholtz1} can be put in bijection with the eigenstates of the
free-particle Hamiltonian $\op H_s$. For positive energies, the quantum numbers of the eigenstates are determined by the Laplacian operator,
regardless of the particular features of the potential.

@@

\begin{figure}
\includegraphics[scale=0.6]{Fig2_CO_spectrum.eps}
\caption{\label{co_spectrum}(Color online) Illustrative LR absorption spectra of the $CO$ molecule in computational setups with different sizes of the simulation domain. The value of the IP threshold is indicated by a vertical dashed line.}
\end{figure}

@@

\section{Computational details} \lb{compdetails}
The illustrative calculations presented in this paper are all KS-DFT calculations at the LDA level of theory. We employed the BigDFT code \cite{BigDFT},
which employs Daubechies Wavelets as computational basis set.
Such orthonormal basis presents optimal features of in view of reproducibility of the results,
and it has proven to represent precisely and explicitly system with Free as well as Periodic boundary conditions (BC), without any basis set superposition error nor supercell aliasing, for Free BC.

>>>>

In this paper, we focus our attention to Linear-Response treatments of \emph{molecular} systems,
in particular referring to LR-TDDFT calculation. We will investigate the impact that
\emph{open} (i.e. Isolated) Boundary Conditions (BC) that have to be imposed to the system will have on the computational evaluation of LR quantities.
In order to do this, we perform analytic considerations on the equations beyond LR treatment and discuss the interplay between reproducible LR quantities and the \emph{locality} of the related objects. Such considerations will enable us to distinguish quantities that \textit{a priori} can be compared among different treatments - and therefore tested for reproducibility - from others that would \emph{explicitly} depend of the employed computational treatment.

\added[id=dal]{This analysis evidences that the description of molecular systems beyond the ground state unavoidably leads to the inclusion of delocalized
degrees of freedom above a specific energetic threshold. Moreover on the basis of this analysis we can assess something about the reliability of the results provided by computational setups able to explicit include the presence of delocalized degrees of freedom compared
to the ones in which these degrees of freedom are excluded by design.}

@@

Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of discussing the impact of the BC, it is instructive to transform
the Schr\"odinger equation in a scattering problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \GH(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\GH(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

@@

On the other hand, positive energy solutions behave differently.
Even though the potential is localized \replaced[id=dal]{the associated Helmholtz kernel is a spherical wave and its}{, their positive energy are spherical waves and the} convolution with the source term provides a delocalized wavefunction $\psi(\r)$ in the whole space.
As they cannot be normalized they belong to the essential spectrum of the Hamiltonian operator and form a \emph{continuum} of states.
Outside of the support of the $\ket{\op V \psi}$ the solutions of Eq.~\eqref{SEHelmholtz1} can be put in bijection with the eigenstates of the
free-particle Hamiltonian $\op H_s$. For positive energies, the quantum numbers of the eigenstates are determined by the Laplacian operator,
regardless of the particular features of the potential.

@@

\begin{figure}
\centering
\subfloat
{\includegraphics[scale=0.56]{Fig2a_CO_AlphaOmegaReal.eps}} \\
\centering
\subfloat
{\includegraphics[scale=0.58]{Fig2b_CO_spectrum.eps}}
\caption{\label{co_spectrum}(Color online) LR dynamical polarizability of the $CO$ molecule in computational setups with different 
sizes of the simulation domain. The top panel displays the real part and the bottom one the imaginary part. The value of the IP threshold 
is indicated by a vertical dashed line.}
\end{figure}

% \begin{figure}
% \includegraphics[scale=0.6]{Fig2_CO_spectrum.eps}
% \caption{\label{co_spectrum}(Color online) Illustrative LR absorption spectra of the $CO$ molecule in computational setups with different sizes of the simulation domain. 
% The value of the IP threshold is indicated by a vertical dashed line.}
% \end{figure}

@@

\section{Computational details} \lb{compdetails}
The illustrative calculations presented in this paper are all KS-DFT calculations at the LDA level of theory. We employed the BigDFT code \cite{BigDFT},
that make usage of Daubechies Wavelets as computational basis set. This orthonormal basis presents optimal features of in view of reproducibility of the results, and it has proven to represent precisely and explicitly system with Free as well as Periodic boundary conditions (BC), without any basis set superposition error nor supercell aliasing, for Free BC.
